---
title: "Homework 3"
author: "Michael Weigand"
format: html
editor: visual
[Link to my GitHub Repository](https://github.com/michiwei2002-bit/ENVS-193DS_homework-03.git)
---



```{r}
# Load packages
library(tidyverse)
library(here)
library(janitor)

# Read in the soil salinity data
salinity <- read_csv(here("Data", "salinity-pickleweed.csv")) %>% 
  clean_names()
```


```{r}
# Read in personal data from 'Data' folder
personal_data <- read_csv(here("Data", "personal_pushup_data.csv")) %>% 
  clean_names()
```


#Problem 1: Slough soil salinity

#1a. Selecting the Appropriate Statistical TestThe two appropriate tests to determine the strength of the relationship are Pearson’s Correlation (r) and Simple Linear Regression.

#Pearson’s Correlation (r): This test tells us how closely the data points follow a straight line. It provides a single number between -1 and 1 that represents the direction and strength of the relationship but does not imply that one variable causes the other.

#Simple Linear Regression: This test goes further by creating a mathematical model (a line of best fit) to describe the relationship. It allows us to predict the biomass based on a specific salinity level and tells us how much of the variation in biomass is actually explained by the salinity.



#1b
```{r}
# Create scatterplot to show the relationship between salinity and biomass
ggplot(data = salinity, aes(x = salinity_m_s_cm, y = pickleweed)) +
  geom_point(color = "darkcyan", size = 2) + # Change color from default and adjust size
  theme_bw() +                             # Change theme from default (white background)
  labs(
    x = "Soil Salinity (mS/cm)",           # Relabel x-axis with units
    y = "Pickleweed Biomass (g)",          # Relabel y-axis with units
    title = "Pickleweed Biomass vs. Soil Salinity",
    subtitle = "Analysis of plant growth in brackish slough restoration site"
  )
```


#1c.
```{r}
# Checking for linearity: We look at the scatterplot we made in 1b.
# Checking for normality: We use a Q-Q plot to see if the biomass follows a normal distribution.
ggplot(salinity, aes(sample = pickleweed)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  theme_minimal() +
  labs(title = "Q-Q Plot for Pickleweed Biomass")
```

#I checked the assumptions of linearity and normality for this analysis. I assessed linearity by reviewing the scatterplot created in Part 1b and checked the normality of the biomass data using the Q-Q plot above. Because the points in the Q-Q plot closely follow the straight diagonal line, I assess that the data is approximately normally distributed and meets the assumptions for a Pearson’s correlation.


```{r}
# Running a Pearson's Correlation Test
# This measures the strength and direction of the relationship between salinity and biomass
cor.test(salinity$salinity_m_s_cm, salinity$pickleweed)
```

#To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a Pearson’s correlation test. I found a moderate, positive relationship between the two variables (r = 0.53, p = 0.0086), suggesting that as soil salinity increases, the biomass of the California pickleweed also tends to increase.



#1d.Which test used and why:To evaluate the strength of the relationship between California pickleweed biomass and soil salinity, I used a Pearson’s correlation test because both variables are continuous and numerical, and my assumption checks confirmed the data is approximately normally distributed.

#Interpretation of test:The test reveals a moderate, positive relationship between soil salinity and plant biomass, indicating that pickleweed plants at this restoration site tend to grow larger as the salt concentration in the soil increases (Pearson’s r = 0.53, df = 21, p = 0.0086).



#1e.Our analysis indicates a moderate positive relationship between soil salinity and pickleweed growth, suggesting that these plants are more successful in saltier conditions at our restoration site. To maximize restoration success, we should prioritize planting new pickleweed in areas with higher electrical conductivity, as these locations are currently producing significantly more biomass.


#1f.
```{r}
# Running a Simple Linear Regression
# Format: lm(y_variable ~ x_variable, data = your_data_name)
salinity_model <- lm(pickleweed ~ salinity_m_s_cm, data = salinity)

# Display the summary of the model to see p-values and coefficients
summary(salinity_model)
```

#Both the Pearson’s correlation and the simple linear regression led to the same decision to reject the null hypothesis, as both tests yielded a statistically significant p-value of 0.0086. While the correlation focused on the strength of the linear relationship ($r = 0.53$), the regression model provides a specific slope (Estimate = 2.08), which tells us that for every 1 mS/cm increase in soil salinity, we expect an average increase of approximately 2.08 grams in pickleweed biomass. Because both tests show a significant, positive trend, my interpretation of the biological relationship remains the same: higher salinity supports better growth for pickleweed at this location.
