---
title: "Homework 3"
author: "Michael Weigand"
format: html
editor: visual
[Link to my GitHub Repository](https://github.com/michiwei2002-bit/ENVS-193DS_homework-03.git)
---



```{r}
# Load packages
library(tidyverse)
library(here)
library(janitor)

# Read in the soil salinity data
salinity <- read_csv(here("Data", "salinity-pickleweed.csv")) %>% 
  clean_names()
```


```{r}
# Read in personal data from 'Data' folder
personal_data <- read_csv(here("Data", "personal_pushup_data.csv")) %>% 
  clean_names()
```


#Problem 1: Slough soil salinity

#1a. Selecting the Appropriate Statistical TestThe two appropriate tests to determine the strength of the relationship are Pearson’s Correlation (r) and Simple Linear Regression.

#Pearson’s Correlation (r): This test tells us how closely the data points follow a straight line. It provides a single number between -1 and 1 that represents the direction and strength of the relationship but does not imply that one variable causes the other.

#Simple Linear Regression: This test goes further by creating a mathematical model (a line of best fit) to describe the relationship. It allows us to predict the biomass based on a specific salinity level and tells us how much of the variation in biomass is actually explained by the salinity.



#1b
```{r}
# Create scatterplot to show the relationship between salinity and biomass
ggplot(data = salinity, aes(x = salinity_m_s_cm, y = pickleweed)) +
  geom_point(color = "darkcyan", size = 2) + # Change color from default and adjust size
  theme_bw() +                             # Change theme from default (white background)
  labs(
    x = "Soil Salinity (mS/cm)",           # Relabel x-axis with units
    y = "Pickleweed Biomass (g)",          # Relabel y-axis with units
    title = "Pickleweed Biomass vs. Soil Salinity",
    subtitle = "Analysis of plant growth in brackish slough restoration site"
  )
```


#1c.
```{r}
# Checking for linearity: We look at the scatterplot we made in 1b.
# Checking for normality: We use a Q-Q plot to see if the biomass follows a normal distribution.
ggplot(salinity, aes(sample = pickleweed)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  theme_minimal() +
  labs(title = "Q-Q Plot for Pickleweed Biomass")
```

#I checked the assumptions of linearity and normality for this analysis. I assessed linearity by reviewing the scatterplot created in Part 1b and checked the normality of the biomass data using the Q-Q plot above. Because the points in the Q-Q plot closely follow the straight diagonal line, I assess that the data is approximately normally distributed and meets the assumptions for a Pearson’s correlation.


```{r}
# Running a Pearson's Correlation Test
# This measures the strength and direction of the relationship between salinity and biomass
cor.test(salinity$salinity_m_s_cm, salinity$pickleweed)
```

#To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a Pearson’s correlation test. I found a moderate, positive relationship between the two variables (r = 0.53, p = 0.0086), suggesting that as soil salinity increases, the biomass of the California pickleweed also tends to increase.



#1d.Which test used and why:To evaluate the strength of the relationship between California pickleweed biomass and soil salinity, I used a Pearson’s correlation test because both variables are continuous and numerical, and my assumption checks confirmed the data is approximately normally distributed.

#Interpretation of test:The test reveals a moderate, positive relationship between soil salinity and plant biomass, indicating that pickleweed plants at this restoration site tend to grow larger as the salt concentration in the soil increases (Pearson’s r = 0.53, df = 21, p = 0.0086).
